---
title: "Overview"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidymodels)
```

[Tidymodels](https://www.tidymodels.org/) is a collection of packages developed for modeling and machine learning. These packages simplify the task of fitting a model to data by providing tools to assist with data pre-proprocessing, model initiation, and results valiation. Tidymodels emphasizes simplicity and readability by abstracting complex modelling operations into common interfaces and optimizing functions for the [pipe operator](https://towardsdatascience.com/an-introduction-to-the-pipe-in-r-823090760d64).

To use Tidymodels, simply install and load the `tidymodels` package. The `tidymodels` package loads all of its core modeling packages as well as some helpful `tidyverse` packages such as `dplyr` and `ggplot2`. For a complete list of packages loaded by Tidymodels, reference their [packages page](https://www.tidymodels.org/packages/).

```{r, eval=FALSE}
library(tidymodels)
```

Each tab on this site corresponds to a different package from Tidymodels: 

* [`RSample`](RSample.html) - Re-sample and split training/testing data
* [`Recipes`](Recipes.html) - Create "pre-processing objects" that apply transformations to data 
* [`Parsnip`](Parsnip.html) - Common, simple interface for model initiation
* [`Yardstick`](Yardstick.html) - Evaluate model performance
* [`Workflows`](Workflows.html) - Bundle recipes and models into a single object
* [`Workflowsets`](WorkflowSets.html) - Create and compare sets of workflows
* [`Tune`](Tune.html) - Tune hyperparameters for models 

Additionally, this page walks through some examples of modeling with tidymodels:

* [Iris](Iris.html) - Predicting Iris species from measurements.

<br>

### Dataset ###

This site introduces the Tidymodels packages by walking through the complete modelling process with an example dataset, defined below. 

`data_iris` is a modified version of the inbuilt `iris` dataset that includes row number as a variable. 
```{r}
data_iris <- bind_cols(all_of(iris), 
                       row = 1:nrow(iris))

head(data_iris)
```

Using this example dataset, we will predict `Species` based on the other variables (row is an identification variable and not used as a predictor). This modelling exercise spans across multiple pages on this site -- the complete code is aggregated at the [Iris modelling example](Iris.html). 

<br>

### General Resources ###

* https://rviews.rstudio.com/2019/06/19/a-gentle-intro-to-tidymodels/ 
  * A simple walkthrough of some core Tidymodels packages and concepts. 
* https://www.tmwr.org/
  * A highly detailed guidebook of Tidymodels and how to create accurate models. 
  
<br>
  
### About ###

This site is authored by Omi Johnson and is meant to serve as an introductory walkthrough of Tidymodels. This site was built using RMarkdown and GitHub hosting -- find a simple walkthrough of how to build your own RMarkdown site [here](https://psyteachr.github.io/hack-your-data/rmarkdown_2.html). The source repository for this site is on [my GitHub](https://github.com/oj713/tidymodels).

**NOTE**: This is not a comprehensive guide. These notes are incomplete and only mention methods/techniques that the author is familiar with. Descriptions of packages and methods most likely do not include all possible arguments or applications. For a complete list of each package's methods and capabilities, please look at the official documentation. 