---
title: "Overview"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

https://www.tidymodels.org/ 

Tidymodels is a collection of packages developed for modeling and machine learning. These packages simplify the task of fitting a model to data by providing tools to assist with data pre-proprocessing, model initiation, and results valiation.

To use Tidymodels, simply install and load the `tidymodels` package. The `tidymodels` package loads all of its core modeling packages as well as some helpful `tidyverse` packages such as `dplyr` and `ggplot2`. 

```{r, eval=FALSE}
library(tidymodels)
```

Each tab on this site corresponds to a different package from Tidymodels: 

* `RSample` - Re-sample and split training/testing data
* `Recipes` - Create "pre-processing objects" that apply transformations to data 
* `Parsnip` - Common, simple interface for model initiation
* `Workflows` - Bundle recipes and models into a single object
* `Yardstick` - Evaluate model performance

<br>

### General Resources ###

* https://rviews.rstudio.com/2019/06/19/a-gentle-intro-to-tidymodels/ 
  * A simple walkthrough of some core Tidymodels packages and concepts. 
* https://www.tmwr.org/
  * A detailed guidebook of Tidymodels and general modeling concepts 
  
<br>
  
### About ###

This site is authored by Omi Johnson and is meant to serve as an introductory guide/walkthrough of Tidymodels. 

**NOTE**: This is not a comprehensive guide. These notes are incomplete and only mention methods/techniques that the author is familiar with. Descriptions of packages and methods most likely do not include all possible arguments or applications. For a complete list of each package's methods and capabilities, please look at the official documentation. 