---
title: "Welcome!"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

[Tidymodels](https://www.tidymodels.org/) is a collection of packages developed for modeling and machine learning. These packages simplify the task of fitting a model to data by providing tools to assist with data pre-proprocessing, model initiation, and results valiation. Tidymodels emphasizes simple, readable code by abstracting complex modelling operations into common interfaces and optimizing functions for the [pipe operator](https://towardsdatascience.com/an-introduction-to-the-pipe-in-r-823090760d64).

To use Tidymodels, simply install and load the `tidymodels` package. The `tidymodels` package loads all of its core modeling packages as well as some helpful `tidyverse` packages such as `dplyr` and `ggplot2`. Find a complete list of Tidymodels packages [here](https://www.tidymodels.org/packages/).  

```{r, eval=FALSE}
library(tidymodels)
```

<br>

## Overview ##

Each tab on this site corresponds to a different package from Tidymodels: 

* [`RSample`](RSample.html) - Re-sample and split training/testing data
* [`Recipes`](Recipes.html) - Create "pre-processing objects" that apply transformations to data 
* [`Parsnip`](Parsnip.html) - Common, simple interface for model initiation
* [`Yardstick`](Yardstick.html) - Evaluate model performance
* [`Workflows`](Workflows.html) - Bundle recipes and models into a single object
* [`Tune`](Tune.html) - Tune hyperparameters for models 
* [`Workflowsets`](WorkflowSets.html) - Create and compare sets of workflows

These pages are meant to be read in order, as modelling examples are built cumulatively. To view abbreviated versions of the code through each tutorial, view the following tabs: 

* [Iris](Iris.html) - Predicting Iris species from their measurements.
* [Cfin](Cfin.html) - Optimizing and comparing models of *C. Finmarchicus* abundance. 

<br>

## General Resources ##

* https://rviews.rstudio.com/2019/06/19/a-gentle-intro-to-tidymodels/ 
  * A simple walkthrough of some core Tidymodels packages and concepts. 
* https://www.tmwr.org/
  * A highly detailed guidebook of Tidymodels and how to create accurate models. Much of this tutorial site is just a simplified rehash of information from this guidebook -- if you want to take a deep dive into tuning and optimizing your model, I highly recommend going through this resource.
  
<br>
  
## About ##

This site is authored by Omi Johnson and is meant to serve as an introductory walkthrough of Tidymodels. This site was built using RMarkdown and GitHub hosting -- find a simple walkthrough of how to build your own RMarkdown site [here](https://psyteachr.github.io/hack-your-data/rmarkdown_2.html). The source repository for this site is on my [GitHub](https://github.com/oj713/tidymodels).

Questions? Comments? Contact me by email at **ojohnson@bigelow.org**

**NOTE**: This is not a comprehensive guide. These notes are incomplete and only mention methods/techniques that the author is familiar with. Descriptions of packages and methods most likely do not include all possible arguments or applications. For a complete list of each package's methods and capabilities, please look at the official documentation. 