---
title: "Parsnip"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source("iris_setup.R")
```

When building a model for your dataset, there are a wide variety of models and model packages that each have their own unique parameters and specifications. [`Parsnip`](https://parsnip.tidymodels.org/) provides a unified, simple interface for model creation so that you can easily swap between models without worrying about minor syntax changes. 

<br>

### Initializing a Model ### 

A `parsnip` model is built by defining an initial model object, setting its parameters/engine, and then fitting the model to data. Each type of supported model has its own initializing function. 

For example, suppose we wanted to create a linear regression model. To initiate this model, call the [`linear_reg()`](https://parsnip.tidymodels.org/reference/linear_reg.html) function. 

```{r} 
lr <- linear_reg()
```

Alternatively, to initialize a random forest model use the [`rand_forest()`](https://parsnip.tidymodels.org/reference/rand_forest.html) function. 

```{r}
rf <- rand_forest()
```

Other models supported include logistic regression, boosted trees, k nearest neighbors, and more. For a full list of supported models see the official [Function Reference](https://parsnip.tidymodels.org/reference/index.html).

<br>

### Setting Parameters ###

After initializing a model, set its specifications/parameters. 

* [**`set_mode()`**](https://parsnip.tidymodels.org/reference/set_args.html) determines whether the model is a classification or regression model. Not every model has both options -- for example, `linear_reg()` only supports regression. 

* [**`set_args()`**](https://parsnip.tidymodels.org/reference/set_args.html) sets parameters specific to the model type. The names of these arguments are standardized so that similar parameters across models have the same name -- for example, `penalty` is the name of the regularization argument, if a model has one. 

```{r modeargs}
# setting mode and parameters for both of our example models 
# note that you don't need to set every model parameter if you're happy with the defaults
lr <- lr |>
  set_mode("regression") |> 
  set_args(penalty = NULL) 

rf <- rf |>
  set_mode("classification") |>
  set_args(trees = 200)
```

* [**`set_engine()`**](https://parsnip.tidymodels.org/reference/set_engine.html) determines the model *engine*, or package that will handle model creation. For many models, there are multiple packages that can run that model, each with their own unique syntax and interface. `parsnip` unifies the interface, but doesn't replace the modeling package -- meaning that the underlying modelling engine must be specified. To switch engines later on, simply change the argument for `set_engine()`. 

  * Engines might require package installs. 
  * use [`show_engines()`](https://parsnip.tidymodels.org/reference/show_engines.html) retrieve a list of all supported engines for a model. Some engines require a `parsnip` extension package and may not be shown. 
  * To see how `parsnip` arguments transfer to a specific engine, use the [`translate()`](https://parsnip.tidymodels.org/reference/translate.html) method. 
  * Some models have engine-specific arguments that do not translate to other engines. Pass these arguments to `set_engine()`. 

```{r engines}
# what engines are available for random forest?
show_engines("rand_forest")

# setting the engine to randomForest 
# note the engine-specific argument used in this call 
rf <- rf |>
  set_engine("ranger", verbose = TRUE)

# to change the engine, call set_engine() again
rf <- rf |>
  set_engine("randomForest")

# using translate() to view the model and how parsnip arguments transfer to the randomForest package
translate(rf)
```

**Note**: Mode, engine, and model arguments can be specified within the initial model call. However, setting them with separate functions enhances readability and creates a more flexible interface.

* Engine-specific arguments must be declared within `set_engine()`. 

```{r}
# defining parameters within initial call
rand_forest(mode = "classification", 
            engine = "randomForest", 
            trees = 500)
```

<br>

### Fitting ###

Once a model has been initialized and defined, execute the model on our prepped training data using the `fit()` function. `fit()` takes in a formula and a model object and returns a fitted model. 

We will use the `iris_recipe_prepped`, `prepped_training` and `prepped_testing` datasets defined in the [Recipes Tutorial](Recipes.html). 

```{r}
# extracting model formula from prepped recipe object
rf_formula <- formula(iris_recipe_prepped)

# fitting our random forest model to our training data
rf_fit <- fit(rf, rf_formula, data = prepped_training)

# taking a look at our fitted model
rf_fit 
```

**Note:** If a model uses a specialized formula that specifies model structure as well as terms (eg. GAMs), pass it to `fit()`. 

**Note:** If the recipe used to preprocess data modified any roles from your original formula, be sure to extract your model formula from the recipe object to account for those changes. 

<br>

### Predicting ###

In order to generate predictions from new data, pass the fitted model and a dataset to either the [`predict()`](https://parsnip.tidymodels.org/reference/predict.model_fit.html) or [`augment()`](https://parsnip.tidymodels.org/reference/augment.html) methods `predict()` will return a tibble of predictions, and `augment()` will bind prediction columns to the input data. 

* Predictions are returned as a tibble with predictable column names.
* `predict()` has a `type` argument used to specify prediction type. Options for `type` are "numeric", "class", "prob", "conf_int", "pred_int", "quantile", "time", "hazard", "survival", or "raw". `type` defaults to either "numeric" or "class" depending on model mode. Not all `type` options will work for every model. 


```{r}
# generating predictions
predict(rf_fit, prepped_testing) |>
  head()

# note the use of type = prob to retrieve probabilities for each class
predict(rf_fit, prepped_testing, type = "prob") |>
  head()

# augment binds predictions to original data
# augment does not have a type argument
iris_preds <- augment(rf_fit, prepped_testing) 

head(iris_preds)
```

Methods to fit resampling sets are covered in the [Tune Tutorial]("Tune.html").

<br>

### Further Resources ###

* https://www.tmwr.org/models.html
  * Excellent overview of `parsnip`, especially regarding how it compares to other modelling packages.
