---
title: "Workflow Sets"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source("cfin_setup.R")
```

When building a model, its often a good idea to compare and contrast different options for preprocessing and model creation. [`Workflowsets`](https://workflowsets.tidymodels.org/) contains tools to build sets of workflows for easy comparison.  

<br>

## Building a Workflow Set ##

In order to build a workflow set, begin by creating a list of preprocessors and a list of models. The preprocessors can be either formulas or recipes. 

* To analyse variable contribution, use the helper [`leave_var_out_formulas()`](https://workflowsets.tidymodels.org/reference/leave_var_out_formulas.html). This method takes in a formula and returns a list of formulas with each predictor left out in turn. 
* Recipes and models can have tunable parameters. 

This example uses the `cfin` dataset introduced in the [Tune Tutorial](Tune.html). 

```{r}
# creating three different recipes -- one with no transformations, one where bathymetry is log scaled, and one where bathymetry is log scaled and all variables are normalized. 
basic_rec <- recipe(abundance ~ ., data = cfin_train) |>
  update_role(lat, lon, year, new_role = "ID") |>
  step_corr(threshold = .9) |>
  # boosted regression trees require dummy variables
  step_dummy(all_nominal_predictors())

log_bathy <- basic_rec |>
  step_log(Bathy_depth, base = 10)

normalize <- log_bathy |>
  step_normalize(all_numeric_predictors())

# the names of each item in the list will be later used to id workflows 
preproc_list <- list(basic = basic_rec, 
                     log = log_bathy, 
                     norm = normalize)

# creating two different models -- random forest and boosted regression tree
rf <- rand_forest(mode = "regression", 
                  engine = "ranger", 
                  trees = 100)

# note required package install: xgboost
brt <- boost_tree(mode = "regression", 
                  engine = "xgboost",
                  trees = 15)

model_list <- list(rf = rf, brt = brt)
```

To build the actual set of workflows, use the  [`workflow_set()`](https://workflowsets.tidymodels.org/reference/workflow_set.html) method. Each workflow will combine a preprocessor and recipe from the input lists. 

* Workflow sets can also be created from a list of pre-existing workflows by using the [`as_workflow_set()`](https://workflowsets.tidymodels.org/reference/as_workflow_set.html) method. 

```{r}
cfin_wkfs <- workflow_set(preproc_list, model_list, 
                          cross = TRUE) # create all combinations of models?

cfin_wkfs
```

<br>

## Operating Across Workflows ##

After creating the workflow set, apply operations across the set using the [`workflow_map()`](https://workflowsets.tidymodels.org/reference/workflow_map.html) method. Results are collected into the `results` column of the workflow set. 

* `workflow_map()` can apply any `tune_*()` method as well as `tune::fit_resamples()`.

```{r}
# applying fit_resamples across the workflow set to assess performance
# using cfin_control and cfin_folds defined in Tune Tutorial
fitted_wkfs <- cfin_wkfs |>
  workflow_map(fn = "fit_resamples", # method to apply in quotes
               verbose = FALSE, 
               seed = 400, 
               resamples = cfin_folds, # arguments for fn 
               metrics =  metric_set(rmse, rsq, mae),
               control = cfin_control)

fitted_wkfs
```

<br>

## Analysing Results ##

In order to cleanly assemble results into a tibble, use [`collect_metrics()`](https://workflowsets.tidymodels.org/reference/collect_metrics.workflow_set.html). To determine the best workflow, use [`rank_results()`](https://workflowsets.tidymodels.org/reference/rank_results.html) to sort workflow results by a desired metric. 

```{r}
# collecting metrics
fitted_wkfs |>
  collect_metrics() |>
  head()

# determining the workflow with the best rmse
fitted_wkfs |>
  rank_results(rank_metric = "rmse", select_best = FALSE)
```

`Workflowsets` also contains an [`autoplot()`](https://workflowsets.tidymodels.org/reference/autoplot.workflow_set.html) method to easily visualize results. `autoplot()` is capable of plotting both `rank_results()` as well as the results for individual workflows. 

* When plotting entire workflow set, the x axis is the rank of the workflow. For more illustrative axes use `ggplot` methods on the output of `rank_results()`. 

```{r}
fitted_wkfs |>
  autoplot(rank_metric = "rmse", # metric to determine workflow rank, 
           id = "workflow_set") # workflow to plot -- use workflow_set to plot all
```

**Note:** `Workflowsets` has additional capabilities for editing, extracting, and examining individual workflows that are not covered here. See the "Further Resources" section for more information. 

<br>

## Further Resources ##

* https://www.tmwr.org/tuning.html
  * A much more detailed walkthrough of tuning the model. 
* https://workflowsets.tidymodels.org/articles/tuning-and-comparing-models.html
  * Another walkthrough of workflowsets that covers workflow modification and extraction. 
* https://rpubs.com/marcelocostamagna/730911
  * Contains an example of applying `tune_grid()` to a tunable workflow set. 
