---
title: "Workflows"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source("setup.R")
```

When building a model, preprocessing steps are often specific to a certain model. `workflows` bundle preprocessing and `parsnip` objects together, such that we can prep the data and fit the model with a single call to `fit()`.

<br>

## Creating a Workflow ##

We can initiate a workflow object with the [`workflow()`](https://workflows.tidymodels.org/reference/workflow.html) function. You can add a preprocessor and model in the initial function call or by using `add_` methods (below). 

```{r}
wkf <- workflow(preprocessor = NULL, # pass in preprocessor here 
                spec = NULL) # pass in model here
```

<br>

### Adding a Preprocessor ###

There are two options for a workflow preprocessor: 

* **A Formula or Role Specifications**: for simple models which need no data transformations, the only preprocessing needed is to specify outcome and predictor variables. [`add_formula()`](https://workflows.tidymodels.org/reference/add_formula.html) allows you to pass in a formula, or you can specify outcomes and predictors directly with [`add_variables()`](https://workflows.tidymodels.org/reference/add_variables.html. 

```{r}
wkf1 <- wkf |>
  add_formula(Species ~ .)

# identical to wkf1
# note the use of everything(), which will ignore variables already referenced
wkf |>
  add_variables(outcomes = Species, predictors = everything())
```

* **A Recipe**: for more complex pre-processing, pass in a recipe object using `add_recipe()`. This is more common. For this example, we will use the `iris_recipe` defined in the [Recipes Tutorial](Recipes.html). 
  * You can pass in a prepped or unprepped recipe to the workflow. If the recipe is not prepped, the workflow will prep it for you. 

```{r}
iris_wkf <- wkf |>
  add_recipe(iris_recipe)

iris_wkf
```

Note that each `add_x()` function has accompanying `remove_x()` and `update_x()` functions to allow for workflow modification. 

<br>

### Adding a Model ###

All workflows must include a `parsnip` model object. Add an **unfitted** model object using [`add_model()`](https://workflows.tidymodels.org/reference/add_model.html). We will use the `rf` random forest specification defined in the [Parsnip Tutorial](Parsnip.html).

```{r}
iris_wkf <- iris_wkf |>
  add_model(rf)
```

**Note:** If you are using a model with a specialized formula, add it using the `formula` argument of `add_model()`. You must use `add_model()` for this -- special formulas cannot be specified with the `add_formula()` function. 

```{r, eval = FALSE}
# GAMs have special syntax for formulas because of their smoothing functions
gam <- gen_additive_mod() |>
  set_mode("regression") |>
  set_engine("mgcv")
  
gam_wkf <- workflow() |>
  # notice the lack of special syntax for the preprocessor formula
  add_formula(Sepal.length ~ Species + Sepal.width) |> 
  add_model(gam, 
            # now using GAM specific syntax 
            formula = Sepal.length ~ Species + s(Sepal.width))
```

<br>

## Fitting and Predicting ##

After building a workflow, use [`fit()`](https://workflows.tidymodels.org/reference/fit-workflow.html), [`predict()`](https://workflows.tidymodels.org/reference/predict-workflow.html), and [`augment()`](https://workflows.tidymodels.org/reference/augment.workflow.html) just as you would with a `parsnip` model in order to train the workflow and generate predictions. Input datasets should be raw, **not** baked by a recipe. 

* `augment(<workflow>)` will bind predictions to the unbaked input data. If a recipe contains steps that alter row number, `augment()` will error because the input and output datasets won't have the same length. 
* `fit(<workflow>)` does not take a formula, as the workflow object contains a formula already. 

We will fit and predict the workflow using the `data_split` object defined in the [RSample Tutorial](RSample.html).

```{r}
# fitting the model to the training data
iris_wkf <- iris_wkf |>
  # since we already defined the formula in the workflow, fit() only needs the raw data
  fit(training(data_split)) 

# using augment to generate predictions
# notice that augment now binds predictions to the unbaked input data!
iris_preds <- augment(iris_wkf, testing(data_split))
head(iris_preds)
```

<br>

## Examining Workflows ##

We can view information about a workflow object by calling it. 

```{r}
iris_wkf
```

If we want specific information from a workflow, the `workflows` package contains several extraction methods to retrieve specific aspects of the workflow, such as the fitted model or preprocessor. Several of these methods also work on `parsnip` model objects. 

A complete list of extractors for workflows can be found [here](https://workflows.tidymodels.org/reference/extract-workflow.html).

```{r}
# extracting the (unprepped) preprocessor
extract_preprocessor(iris_wkf)

#extracting the fitted model
extract_fit_parsnip(iris_wkf)
```

<br>

## Further Resources ##

* https://www.tmwr.org/workflows.html
  * Both an introduction of and deep dive into workflows that covers more capabilities than what I have here, including workflow sets. 
