---
title: "Yardstick"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source("setup.R")
```
https://yardstick.tidymodels.org/

The final step of building a model is to evaluate its performance. The `yardstick` package contains various methods to summarize and display model results in a tidy manner. 

We will be evaluating the accuracy of our iris model based on the augmented `iris_preds` data defined in [Workflows Tutorial](Workflows.html). As a note, the predictions have a 100% accuracy rate, so the resulting metrics might not be very illustrative.

<br>

### Metric Methods ###

`yardstick` contains a variety of methods to extract evaluation metrics from a dataset. There are metric methods for both regression and classification models. 

* **Input:** Each `yardstick` metric method takes as argument a dataframe, truth column, and estimate column(s). 
* **Output:** Results will be returned as a tibble containing `metric`, `estimator`, and `estimate` information. 

```{r}
# computing roc_auc and accuracy for iris predictions
roc_auc(iris_preds,
        truth = Species, 
        .pred_setosa:.pred_virginica) # all estimate classes

accuracy(iris_preds, 
         truth = Species, 
         estimate = .pred_class)
```

All metric methods also have a corresponding vectorized method (ending with `_vec`) that simply takes in the truth and estimate vectors and returns the desired metric. 

```{r}
# using the vectoized accuracy metric
accuracy_vec(iris_preds$Species, 
             iris_preds$.pred_class)
```

For a full list of `yardstick` functions, look at the [Official Reference](https://yardstick.tidymodels.org/reference/index.html).

<br>

### Curve Functions ###

`yardstick` contains four curve-based functions: `roc_curve()`, `gain_curve()`, `pr_curve()`, and `lift_curve()`. These methods all have corresponding `ggplot2::autoplot()` methods that allow for easy visualization. 

```{r}
# note that the arguments are the same as those for roc_auc
roc_curve(iris_preds, truth = Species, .pred_setosa:.pred_virginica) |>
  autoplot()
```

### Metric Sets ###

If you want to extract multiple metrics from a dataset, you can package them together into a single method call using [`metric_set()`](https://yardstick.tidymodels.org/reference/metric_set.html). The created method will return each desired metric and its estimate in a tibble.

* Note that metric sets can only contain either classification or regression metrics -- **not** a mix of both. 

```{r}
# say we want to find auc, sensitivity, specificity, and accuracy 
# pass in yardstick method names to metric_set()
my_metrics <- metric_set(roc_auc, sens, spec, accuracy)

# retrieving the metrics using our newly defined method 
# pass in all arguments needed for contained methods
my_metrics(iris_preds, 
           truth = Species,
           estimate = .pred_class, # for sens, spec, and accuracy
           .pred_setosa:.pred_virginica) # for roc_auc
```

The [**`metrics()`**](https://yardstick.tidymodels.org/reference/metrics.html) function is an `yardstick` provided metric set that generates common metrics based on the provided type of input data. For example, for numeric models, `metrics()` will return `rmse`, `rsq`, and `mae`. 

```{r}
# for multiclass models, metrics() will return accuracy and the Kappa statistic
metrics(iris_preds, truth = Species, estimate = .pred_class)
```